---
title: "Signif_custome_category_heatmap"
author: "Kazu"
date: "10/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse);library(cowplot)
```

# custom category heatmap (Fig. 5)
# Making list of custom catetory genes enriched in shade responsive genes (TableS7 for Fig5)
```{r}
## start from this line ## 052616
# fold chagne (shade responsiveness)
load(file.path("..","..","Nozue2016_SAStranscriptome_output","output","summary.vst.response.Rdata") )
dim(summary.vst.response) #[1] 30533    52
head(summary.vst.response)
# find overlap between shade responsive (in Col) and hormone responsive
#test<-read.csv(file.path("..","..","Nozue2016_SAStranscriptome_output","figs_tables","TableS4_hormone.responsiveness.SOMclusters100717.csv") ,row.names=1) # needs to use updated one (092917)
test<-read.csv(file.path("..","..","Nozue2016_SAStranscriptome_output","figs_tables","Dataset2_Custom_enrichment_tsnedbscan_clusters_of_shade-responsive_genes101917v3.csv") ,row.names=1) # tsnedbscan version
test<-test %>% select(-c(Short_description,colE...col.F))
colnames(test)
test<-test %>% filter(gt==1) %>% select(-gt) # select only found in Col clusters. remove gt column

# reading overlapTable
#load(file.path("..","..","Nozue2016_SAStranscriptome_output","output","overlap.GOseq.melt.summary.SOM.customcat.100717.Rdata")) 
overlap.melt.tsnedbscan.customcat.PLoSGenetics2015mutants.errrDE.summary.files<-list.files(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"),pattern="overlap.melt.tsnedbscan.customcat.PLoSGenetics2015mutants.errrDE.summary")

for(y in overlap.melt.tsnedbscan.customcat.PLoSGenetics2015mutants.errrDE.summary.files) 
  { # 140 of y
  load(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",y))
overlap.melt.summary.s<-overlap.melt.summary[!overlap.melt.summary$Var1==0,] # remove cluster "0"
overlap.melt.summary.s<-overlap.melt.summary.s[!overlap.melt.summary.s$Var2=="",]
overlap.melt.summary.s$Var2<-droplevels(overlap.melt.summary.s$Var2)
# select log10>2 & !Var1==0 
overlap.melt.summary.s.significant<-overlap.melt.summary.s[overlap.melt.summary.s$value> -log10(0.01),] # pvalue<0.01 (the same as Fig3)
overlap.melt.summary.s.significant$category.cluster<-with(overlap.melt.summary.s.significant,paste(Var2,Var1,sep="_")) # Var1 = cluster, Var2=category

title<-gsub("(overlap.melt.tsnedbscan.customcat.PLoSGenetics2015mutants.errrDE.summary)(\\.)(eps[[:print:]]+)(\\_)(MinPts[[:digit:]]+)(\\.)(set)(\\.)([[:digit:]]+)(\\.Rdata)","\\3\\4\\5\\6\\7\\9",y)


test.wozero<-test[!test[,title]=="0",c(grep("eps",colnames(test),invert=TRUE,value=TRUE),title)] # remove cluster 0 in test
head(test.wozero)
test.wozero.melt<-melt(test.wozero,id.var=c("AGI",title))
colnames(test.wozero.melt)<-gsub(title,"cluster",colnames(test.wozero.melt))
head(test.wozero.melt)
dim(test.wozero.melt) #6061 4
test.wozero.melt<-test.wozero.melt %>% select(-variable) %>% filter(!value=="")
dim(test.wozero.melt) #191 4
table(test.wozero.melt$AGI) # all Col with "title" 

# # remove small group (less than 3)
#category.morethan3<-names(table(overlap.melt.summary.s.significant$value)[table(test.wozero.melt.hormone.significant$value)>3])
#category.morethan3<-overlap.melt.summary.s.significant %>% count(Var2) %>% filter(n>3) %>% as.data.frame()

test.wozero.melt$category.cluster<-with(test.wozero.melt,paste(value,cluster,sep="_"))
# select (1) category morethan 3 within test.wozero.melt 
#         (2) 
# %>% filter(value %in% category.morethan3$Var2)
test.wozero.melt.hormone.significant<- test.wozero.melt %>% filter(category.cluster %in% overlap.melt.summary.s.significant$category.cluster)
test.wozero.melt.hormone.significant
category.significnat.morethan3<-test.wozero.melt.hormone.significant %>% count(value) %>% filter(n>3) %>% as.data.frame()

if(dim(category.significnat.morethan3)[1]==0) {next} else {

overlap.melt.summary.s.significant$value<-as.character(overlap.melt.summary.s.significant$value)
# test.wozero.melt.hormone.significant.s is my category marker gene list 
test.wozero.melt.hormone.significant.s<-test.wozero.melt.hormone.significant[,c("AGI","value")]
names(test.wozero.melt.hormone.significant.s)<-c("AGI","my.category")
test.wozero.melt.hormone.significant.s.summary<-table(test.wozero.melt.hormone.significant.s)[,-1]
table(test.wozero.melt.hormone.significant.s$my.category)
##

# temp.list<-list()
# for(i in levels(as.factor(test.wozero.melt.hormone.significant.s$AGI))) {
#   temp.list[[i]]<-test.wozero.melt.hormone.significant.s[grep(i,test.wozero.melt.hormone.significant.s$AGI),"my.category"]
# }
######
test.wozero.melt.hormone.significant.s
# reformatting 
test2<-data.frame(AGI=unique(test.wozero.melt.hormone.significant.s$AGI))
my.category<-levels(as.factor(test.wozero.melt.hormone.significant.s$my.category))
for(n in 1:length(my.category)) {
  temp<-rep("",dim(test2)[1])
  names(temp)<-test2$AGI
test.wozero.melt.hormone.significant.s[test.wozero.melt.hormone.significant.s$my.category==my.category[n],"AGI"]
temp[test2$AGI %in% test.wozero.melt.hormone.significant.s[test.wozero.melt.hormone.significant.s$my.category==my.category[n],"AGI"]]<-my.category[n]
temp
test2<-cbind(test2, temp)
}
colnames(test2)[-1]<-my.category
# alternative way
test3<-tidyr::unite(test2, my.category2, colnames(test2)[-1], sep="",remove=TRUE)
category.morethan3.v2<-test3 %>% count(my.category2) %>% filter(n>3) %>% as.data.frame()
test.wozero.melt.hormone.significant.s　%>% count(my.category)
save(test.wozero.melt.hormone.significant.s,file=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",paste("test.wozero.melt.hormone.significant.s.",title,".Rdata",sep="")))
} # category.morethan3 > 0
} # 
```
# comparison of test.wozero.melt.hormone.significant.s gene list (102517)
```{r}
test.wozero.melt.hormone.significant.s.files<-list.files(pattern="test.wozero.melt.hormone.significant.s",path=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"))

 test.wozero.melt.hormone.significant.s.list<- sapply(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",test.wozero.melt.hormone.significant.s.files),function(x) mget(load(x))) # mget 
# rename 
names(test.wozero.melt.hormone.significant.s.list)<-gsub("(../../Nozue2016_SAStranscriptome_output/output/tsne_dbscan_overlapTable_heatmap/test.wozero.melt.hormone.significant.s.)(eps[[:print:]]+)(\\_)(MinPts[[:digit:]]+)(\\.)(set[[:digit:]]+)(.Rdata.test.wozero.melt.hormone.significant.s)","\\2\\3\\4\\5\\6",names(test.wozero.melt.hormone.significant.s.list))
 
test.wozero.melt.hormone.significant.s.list2<-sapply(test.wozero.melt.hormone.significant.s.list,c)
test.wozero.melt.hormone.significant.s.list2["AGI",] # compare this with heatmap (Fig3) for double chekcing
test.wozero.melt.hormone.significant.s.list2["AGI",] %>% sapply(length) %>% tibble() %>% ggplot(aes(.)) + geom_histogram(binwidth=5)

```
# how to look overlap between categories?
```{r}
dim(test.wozero.melt.hormone.significant.s.summary) # 93 10
# pTable
for(i in 2:8) {
  for(n in (i+1):9) {
category.overlapTable<-overlapTable(test.wozero.melt.hormone.significant.s.summary[,i],test.wozero.melt.hormone.significant.s.summary[,n])
category.overlapTable.pTable<-category.overlapTable$pTable
    save(category.overlapTable.pTable,file=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",paste("category.overlapTable.pTable",colnames(test.wozero.melt.hormone.significant.s.summary)[i],colnames(test.wozero.melt.hormone.significant.s.summary)[n],"Rdata",sep=".")))
  }
}
# combine pTable
category.overlapTable.pTable.files<-list.files(pattern="category.overlapTable.pTable",path=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"))

category.overlapTable.pTable.files.list<-sapply(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",category.overlapTable.pTable.files),function(x) mget(load(x))) # mget 
names(category.overlapTable.pTable.files.list)<-gsub("(../../Nozue2016_SAStranscriptome_output/output/tsne_dbscan_overlapTable_heatmap/category.overlapTable.pTable.)([[:print:]]+)(.Rdata.category.overlap)","\\2",names(category.overlapTable.pTable.files.list))
#category.overlapTable.files.DF<-as.data.frame(category.overlapTable.files.list)
p.table<-data.frame(comparison=rep("",length(names(category.overlapTable.files.list))),signif.overlap="")
p.table[1,2]<-category.overlapTable.files.list[[1]][2,2]
p.table[1,1]<-gsub("([[:print:]]+)(Table.pTable)","\\1",names(category.overlapTable.files.list)[1])

for(n in 2:length(names(category.overlapTable.files.list))) {
p.table[n,2]<-category.overlapTable.files.list[[n]][2,2]
p.table[n,1]<-gsub("([[:print:]]+)(Table.pTable)","\\1",names(category.overlapTable.files.list)[n])
}
p.table
# count
for(i in 1:9) {
  for(n in (i+1):10) {
category.overlapTable<-overlapTable(test.wozero.melt.hormone.significant.s.summary[,i],test.wozero.melt.hormone.significant.s.summary[,n])
category.overlapTable.countTable<-category.overlapTable$countTable
    save(category.overlapTable.countTable,file=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",paste("category.overlapTable.count",colnames(test.wozero.melt.hormone.significant.s.summary)[i],colnames(test.wozero.melt.hormone.significant.s.summary)[n],"Rdata",sep=".")))
  }
}
# count
category.overlapTable.count.files<-list.files(pattern="category.overlapTable.count",path=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"))

category.overlapTable.count.files.list<-sapply(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",category.overlapTable.count.files),function(x) mget(load(x))) # mget 
names(category.overlapTable.count.files.list)<-gsub("(../../Nozue2016_SAStranscriptome_output/output/tsne_dbscan_overlapTable_heatmap/category.overlapTable.count.)([[:print:]]+)(.Rdata.category.overlap)","\\2",names(category.overlapTable.count.files.list))
#category.overlapTable.files.DF<-as.data.frame(category.overlapTable.files.list)
count.table<-data.frame(comparison=rep("",length(names(category.overlapTable.count.files.list))),count.overlap="")
count.table[1,2]<-category.overlapTable.count.files.list[[1]][2,2]
count.table[1,1]<-gsub("([[:print:]]+)(Table.countTable)","\\1",names(category.overlapTable.count.files.list)[1])

for(n in 2:length(names(category.overlapTable.count.files.list))) {
count.table[n,2]<-category.overlapTable.count.files.list[[n]][2,2]
count.table[n,1]<-gsub("([[:print:]]+)(Table.countTable)","\\1",names(category.overlapTable.count.files.list)[n])
}
count.table

#### combine signif.overlap and count.overlap
p.table.count<-merge(p.table,count.table,by="comparison")
p.table.count$x<-gsub("([[:print:]]+)(\\.)([[:print:]]+)","\\1",p.table.count$comparison)
p.table.count$y<-gsub("([[:print:]]+)(\\.)([[:print:]]+)","\\3",p.table.count$comparison)
# contingent table format
p.table.count.reshape<-reshape(data=p.table.count[,c("x","y","count.overlap")],idvar="x",timevar="y",direction="wide")
colnames(p.table.count.reshape)<-gsub("count.overlap.","",colnames(p.table.count.reshape))
p.table.count.reshape
p.table.count.reshape[is.na(p.table.count.reshape)]<-""
p.table.count.reshape

######## under construction ######### 102417
# 
#write.csv(test.wozero.melt.hormone.significant.ssss,file="../../Nozue2016_SAStranscriptome_output/figs_tables/TableS6_test.wozero.melt.hormone.significant.102417.csv") # This should be TableS6 (101017)
```
## draw expression pattern of custom categories (significantly enriched); Fig5A
# visualize hormone pathway related genes (1213-1214,2015). Updated (052616-)
```{r}
test.wozero.melt.hormone.significant.s.files<-list.files(pattern="test.wozero.melt.hormone.significant.s",path=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"))

 test.wozero.melt.hormone.significant.s.list<- sapply(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",test.wozero.melt.hormone.significant.s.files),function(x) mget(load(x))) # mget 
# rename 
names(test.wozero.melt.hormone.significant.s.list)<-gsub("(../../Nozue2016_SAStranscriptome_output/output/tsne_dbscan_overlapTable_heatmap/test.wozero.melt.hormone.significant.s.)(eps[[:print:]]+)(\\_)(MinPts[[:digit:]]+)(\\.)(set[[:digit:]]+)(.Rdata.test.wozero.melt.hormone.significant.s)","\\2\\3\\4\\5\\6",names(test.wozero.melt.hormone.significant.s.list))

# summary.vst.response.kazu
load(file.path("..","..","Nozue2016_SAStranscriptome_output","output","summary.vst.response.Rdata")) # response (092917)
summary.vst.response.kazu$AGI <- row.names(summary.vst.response.kazu)
response.melt <- melt(summary.vst.response.kazu,id.vars= "AGI",variable.name = "sample", value.name = "log2.dif") # this value is not log transformed. This is just ratio (responsiveness to shade; 101117)
hist(response.melt$log2.dif)
response.melt$log2.dif<-log2(response.melt$log2.dif)
head(response.melt)
response.melt$gt <- as.factor(regmatches(response.melt$sample,regexpr("^[0-9]{1,2}",response.melt$sample)))
response.melt$time <- as.factor(regmatches(response.melt$sample,regexpr("[0-9]{1,2}hr",response.melt$sample)))
response.melt$rep <- as.factor(regmatches(response.melt$sample,regexpr("[A-Z]$",response.melt$sample)))
head(response.melt)
tail(response.melt)
manuscript.genotypes <- as.character(c(1,3,4,5,6,7,8,9,10,13,14,17,19) )
response.melt <- response.melt[response.melt$gt %in% manuscript.genotypes,]
response.melt <- droplevels(response.melt) # droplevels is useful.
summary(response.melt)

# making loop or function for drawing Fig5 graph
for(n in 21:length(names(test.wozero.melt.hormone.significant.s.list))) {

my.category.diff.heatmap.graph<-list()
#categories <- read.csv(file.path("..","..","Nozue2016_SAStranscriptome_output","figs_tables","TableS6_test.wozero.melt.hormone.significant.ssss.101017.csv"),row.names=1,as.is=TRUE) #  re-normalize data after subsetting in edgeR DE analysis. specific to each category (101017)
categories<-test.wozero.melt.hormone.significant.s.list[[n]]
print(categories)
title<-names(test.wozero.melt.hormone.significant.s.list)[n]
print(title)
head(categories)
table(categories$my.category)
# select categories larger than three genes
categories %>% count(my.category) %>% filter(n>3) %>% as.data.frame() -> selected.category
categories<-categories[categories$my.category %in% selected.category$my.category,]
categories$my.category<-factor(categories$my.category) # categories$my.category has to be factor

#categories$my.category<-factor(categories$my.category,levels=rev(c("IAAup","BLup","BLdown","PIFtarget","ACCup","ABAup","ABAdown","MJup","MJdown","SAup_24h"))) # 120116
#categories$my.category<-factor(categories$my.category,levels=rev(c("IAAup","MJup","ABAdown"))) # 120116
# # summary.vst.response.kazu
# load(file.path("..","..","Nozue2016_SAStranscriptome_output","output","summary.vst.response.Rdata")) # response (092917)
summary.vst.response.kazu<-summary.vst.response
dim(summary.vst.response.kazu) # [1] 30533    52
head(summary.vst.response.kazu)
# for Fig5A, Col
my.category.heatmap.graph<-list()
for(i in c(1,17,19,3,4,5,14,10,6,13,7,8,9)) { # 
my.category.heatmap.graph[[i]]<-my.category.heatmap6(summary3.response=summary.vst.response.kazu,gt.num=i,gt2=conversion.table[conversion.table$num==i,2],newdata.Col.wGO.selected=categories,plotname="my.category.expression.pattern.102617v1.png",w=5,h=5*4/3, path=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"),save.plot=F,legend.fill=F,file.name=title,category.level=levels(categories$my.category)) #  legend.fill=F to T for new Fig5 (all absolute value, previously called FigS2)
}
# summerize data in one data frame
FigS2.absolute.summary.table.melt.files<-list.files(path=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap"),pattern="FigS2.absolute.summary.table.melt") # (102617)
FigS2.absolute.summary.table.melt.files<-grep(paste(title,".Rdata",sep=""),FigS2.absolute.summary.table.melt.files,value=T)
FigS2.absolute.summary.table.melt.list2<-sapply(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",FigS2.absolute.summary.table.melt.files),function(x) mget(load(x))) # mget will return the value of the object(or objects) in a list. see 
FigS2.absolute.summary.table.melt.DF<-do.call(cbind,FigS2.absolute.summary.table.melt.list2)
# remove redundant columns 
colnames(FigS2.absolute.summary.table.melt.DF)<-gsub("(../../Nozue2016_SAStranscriptome_output/output/tsne_dbscan_overlapTable_heatmap/FigS2.absolute.summary.table.melt)([[:print:]]+)(.Rdata.summary.table.melt.)(my.category|variable|value)","\\2.\\4",colnames(FigS2.absolute.summary.table.melt.DF))
FigS2.absolute.summary.table.melt.DF<-FigS2.absolute.summary.table.melt.DF[,c(paste("1",title,".my.category",sep=""),paste("1",title,".variable",sep="") ,grep("value",colnames(FigS2.absolute.summary.table.melt.DF),value=TRUE))]
write.csv(FigS2.absolute.summary.table.melt.DF,file=file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",paste("FigS2.absolute.summary.table.melt.DF",title,".csv",sep=""))) 

# significance calculation (under construction)
summary(categories)
table(categories$my.category)
# select genes in categories
response.melt.merge <- merge(response.melt,categories,by="AGI",all=FALSE)
response.melt.merge$gt <- relevel(response.melt.merge$gt,ref="1")
table(response.melt.merge$gt,response.melt.merge$time)
mutants <- levels(response.melt.merge$gt)[-1] #1 is Col
category.names<- unique(response.melt.merge$my.category)
results <- lapply(mutants,function(gt) {
  sapply(category.names,function(cat.name) {
    tmp <- response.melt.merge[(response.melt.merge$gt=="1" | response.melt.merge$gt==gt) & response.melt.merge$my.category==cat.name,]
    lmer1 <- lmer(log2.dif ~ gt*time + (1|AGI), data=tmp)
    tmp.results <- anova(lmer1)[, "Pr(>F)"]
    print(row.names(anova(lmer1)))
    names(tmp.results) <- row.names(anova(lmer1))
    min(tmp.results[-2]) # we do not care about time
    print(min(tmp.results[-2]))
  })
} )
names(results) <- mutants
results
results.df <- ldply(results,.id="gt")
results.df
colnames(results.df)[-1]<-as.character(category.names)
results.df
# calclate adjusted p-value
results.df.adj <- matrix(
  p.adjust(unlist(results.df[,-1]),"fdr",n=2*length(unlist(results.df[,-1]))), #multiply by 2 because we took the min value of gt and gt*time
  nrow=nrow(results.df))
results.df.adj <- cbind(as.character(results.df[,1]),as.data.frame(results.df.adj))
colnames(results.df.adj) <- colnames(results.df)
results.df.adj

# write.csv(results.df.adj,file.path("..","..","Nozue2016_SAStranscriptome_output","figs_tables","Fig5B_response_p_values.v6.log2.csv")) #101017 three categories
# 
# add p value in heatmap
# response.p.values<-read.csv(file.path("..","..","Nozue2016_SAStranscriptome_output","figs_tables","Fig5B_response_p_values.v6.log2.csv"),row.names=1) # from modified Julin's script (copied from Kazu_test_v2.R) (100717)

response.p.values<-results.df.adj # due to skipping writing csv (102617)
response.p.values.melt<-melt(response.p.values,id="gt") #
#my.categories<-grep("gt",colnames(response.p.values),invert=TRUE,value=TRUE)

# add significant to absolute value heatmap
for(i in c(17,19,3,4,5,14,10,6,13,7,8,9)) {
significant.position<-data.frame(category=levels(categories$my.category),x=0,y=1:length(levels(categories$my.category)),symbol.gt="",symbol.gt.time="")
## uncer construction. order of my.categories in significant.position should be the same as levels of category in heatmap.
## use category.level paramenter in heatmap function (102617)
response.p.values.melt.gt<-response.p.values.melt[response.p.values.melt$gt==i,]
significant.position$symbol.gt<-(as.numeric(as.character(vlookup(significant.position$category,response.p.values.melt.gt[,c(2,3)],2,F)))<0.01)*1 # significant level
  significant.position$symbol.gt<-gsub("1","*",significant.position$symbol.gt)
  significant.position$symbol.gt<-gsub("0","",significant.position$symbol.gt)

# add significance to heatmap
my.category.heatmap.graph[[i]]<-my.category.heatmap.graph[[i]] + geom_text(data=significant.position,aes(x=x-0.2,y=y,label=symbol.gt),size=20,color="red") #, remove lines for x and y axis. "gt" is red.
rm(significant.position)
}

# supplement Fig2 (absolute value), which will become new Fig5 (101117)
# extract legend
# since Col with legend is different in size from others. I need to extract legend and plot it in different space.
# see http://stackoverflow.com/questions/12041042/how-to-plot-just-the-legends-in-ggplot2
# ibrary(gridExtra)
# g_legend<-function(a.gplot){ 
#   tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
#   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
#   legend <- tmp$grobs[[leg]] 
#   return(legend)} 
# legend <- g_legend(my.category.heatmap.graph[[1]])  # does not work (102617)
#
my.category.heatmap.all <- plot_grid(my.category.heatmap.graph[[1]]+theme(legend.position="none"), 
                                     ggplot()+geom_blank(), ggplot()+geom_blank(),my.category.heatmap.graph[[17]],my.category.heatmap.graph[[19]],my.category.heatmap.graph[[3]],my.category.heatmap.graph[[4]],my.category.heatmap.graph[[5]],my.category.heatmap.graph[[14]],my.category.heatmap.graph[[10]],my.category.heatmap.graph[[6]],my.category.heatmap.graph[[13]],my.category.heatmap.graph[[7]],my.category.heatmap.graph[[8]],my.category.heatmap.graph[[9]],
                      labels=c("A","","","B","","","","","","","","","","",""),
                      vjust = 1.1, hjust=-0.5,
                      ncol=3,nrow=5,
                      label_size = 50)
save_plot(file.path("..","..","Nozue2016_SAStranscriptome_output","output","tsne_dbscan_overlapTable_heatmap",paste("Fig5_oldFig_S2_all.my.category.expression.pattern.v16.log2.signif.",title,".png",sep="")), my.category.heatmap.all,
          ncol = 3, # we're saving a grid plot of 2 columns
          nrow = 5, # and 2 rows
          base_aspect_ratio = 2.5,
          base_height=6,
          limitsize = FALSE
)

} # loop for test.wozero.melt.hormone.significant.s.list
```
# the end of custom category heatmap (shade responsiveness) (Fig. 5)
